---
title: "excise report"
author: "Krisna Gupta"
format:
  html:
    code-fold: true
  pdf:
    mainfont: serif
    documentclass: scrreprt
    papersize: a4
  docx: default
---

## Method

For the aggregate traditional and electric, we use weighted average.

## Data

Some problems with the data:

1. HJE and HTP only available for 2022 and 2023 June.

## Elasticity

Let a standard elasticity parameters

$$
Q_i=AP_i^{-\varepsilon}
$$

which can be approximately log-linearized

$$
q_i=a-\varepsilon p_i
$$

where a lowercase is the log version of its uppercase counterparts. We can, thus, econometrically estimated the above equation with a regression. Additionally, we follow the theory in having an income elasticity with GDP per capita as a proxy. We then estimate:

$$
q_{it}=a-\varepsilon p_{it} + \gamma y_{it} + \epsilon_{it}
$$

We assume an iid $\epsilon_{it}$ for now and uses own-price elasticity since we lack information on the price of electric cigs. The parameter $\varepsilon$ is the own-price elasticity of demand, which we expect to be negative, while $\gamma$ is the income elasticity of demand which is assumed to be positive.

```{r}
#| echo: false
library(tidyverse)
library(readxl)

dat<-read_excel("cukai_use.xlsx",sheet="gab")
dat2<-read_excel("cukai_use.xlsx",sheet="gabb")

plot(dat$qtrad,dat$ptrad,xlab="log quantity",ylab="log price")
plot(dat$revtrad,dat$ttrad,xlab="log revenue",ylab="tariff rate")

```

### Elasticity

```{r}
#| echo: false
reg<-lm(qtrad~ptrad+y,data=dat)
summary(reg)
```

### Revenue from price and from tax rate

```{r}
#| echo: false
reg2<-lm(revtrad~ptrad+y,data=dat)
summary(reg2)
reg3<-lm(revtrad~ttrad+y,data=dat)
summary(reg3)
```

```{r}
#| echo: false
library(vars)
library(ARDL)
dat3<-dat2 |> subset(select=c(revtrad,ttrad))
wew<-ardl(data=dat3,revtrad~ttrad,order = c(1,0))
```

